---
title: "Homework 2"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

First, create two vectors, one corresponding to pre-treatment blood pressure values and the other to post-treatment blood pressure values. 

```{r}
    pre_bp<-c(120,151,125,126,115,132,132,129,134,139,127
              ,122,127,135,133,128,147,138,140,132)
    post_bp<-c(127,145,135,122,115,122,123,126,126,129,132
               ,146,120,114,121,120,128,120,133,115)
```

Using names() and paste(), the elements in both vectors are given the names "Subject_(Subject Number)". This keeps the matched pairs (before and after measurements for each patient) linked.

```{r}
    names(pre_bp)<-paste("Subject",1:20,sep="_")
    names(post_bp)<-paste("Subject",1:20,sep="_")

```

To calculate the difference between pre-treatment blood pressure and post-treatment blood pressure for each subject, the post_bp vector is subtracted from the pre_bp vector. The vector of differences is called diff_bp.

```{r}
    diff_bp<- pre_bp - post_bp
```

Using mean(), the average decrease in blood pressure is calculated.

```{r}
    mean(diff_bp)
```

The patients that experienced a decrease in blood pressure are determined using which(), as those patients have positive values in diff_bp.

```{r}
    which(diff_bp > 0, useNames=TRUE)
```

To create a vector containing only the differences for the above patients, diff_bp is subset by pos_diff_subjects, a vector containing the name of each element that qualifies.

```{r}
    pos_diff_subjects<-c("Subject_2","Subject_4","Subject_6", "Subject_7",
                         "Subject_8","Subject_9","Subject_10","Subject_13",
                         "Subject_14","Subject_15","Subject_16","Subject_17",
                         "Subject_18","Subject_19","Subject_20")
    pos_diff_bp<-diff_bp[pos_diff_subjects]
```

The average decrease in blood pressure is calculated using mean() from the vector created above, pos_diff_bp.

```{r}
    mean(pos_diff_bp)
```

## Task 2: Basic Data Frame Practice

A data frame is created with four variables from four vectors. The patient vector is created below, and pre_bp, post_bp, and diff_bp are called from above.

```{r}
    patient<-paste(1:20,sep="_")
    my_bp<-data.frame(patient,pre_bp,post_bp,diff_bp)
```

After viewing the data frame, the rows for the patients with negative diff_bp values are returned.

```{r}
    my_bp[c(1,3,11,12),]
```

A new variable is added to the data frame, post_bp_less_than_120, that indicates through TRUE/FALSE if the post-treatment blood pressure is less than 120. The new variable was created using ifelse() and the data frame is printed below.

```{r}
    my_bp$post_bp_less_than_120<-ifelse(my_bp$post_bp<120,"TRUE","FALSE")
    knitr::kable(my_bp)
```

## Task 3: List Practice

A new data frame for a new set of 10 patients is created like the one in Task 2, with all 5 variables.

```{r}
    patients<-paste(1:10,sep="_")
    placebo_pre_bp<-c(138,135,147,117,152,134,114,121,131,130)
    placebo_post_bp<-c(105,136,123,130,134,143,135,139,120,124)
    names(placebo_pre_bp)<-paste("Subject",1:10,sep="_")
    names(placebo_post_bp)<-paste("Subject",1:10,sep="_")
    placebo_diff_bp<- placebo_pre_bp - placebo_post_bp
    new_bp<-data.frame(patients,placebo_pre_bp,placebo_post_bp,
                       placebo_diff_bp)
    new_bp$post_bp_less_than_120<-ifelse(new_bp$placebo_post_bp<120,"TRUE","FALSE")
```

The data frame from Task 2, called treatment, and the new data frame created above, called placebo, are stored as two elements in a list.

```{r}
    my_list<-list(treatment=my_bp,placebo=new_bp)
```

There are multiple different ways to access one element of the list. The first two methods shown call the data frame for the first element, while the third method calls the list for the first element.

```{r}
    my_list[[1]]
    my_list$treatment
    str(my_list[1])
```

The pre_bp column of the placebo data is accessed by calling the second column of the second element.

```{r}
    my_list[[2]][2]
```